---
title: "munya-m"
format: html
---

## Research Question:
How do early and mid-career pay levels vary across U.S. states, and what roles do STEM enrollment and alumni perception (making the world better) play in this variation?

## Why this is important?
Career Planning for Students:
students can make informed decisions about which schools or states offer better long-term financial prospects Especially relevant for students deciding between STEM and non-STEM paths.

Education Policy and Funding:
It highlights states or institutions that produce high-earning graduates, which could inform public investment or scholarship programs.

STEM and Social Impact Balance:
It examines whether higher STEM enrollment aligns with better salary outcomes — and whether that comes at the expense of perceived social impact; promoting discussion on balancing career success with societal contributions.

Economic Mobility and Equality:
Understanding salary outcomes across states helps assess regional disparities in economic opportunity.It may inform state-level efforts to improve education outcomes and workforce development.



```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(readr)
```

```{r}
salary <- read_csv("../appx/salary_potential.csv")

```



```{r}
# early career vs mid career pay
ggplot(salary, aes(x = early_career_pay, y = mid_career_pay)) +
  geom_point(alpha = 0.6, color = "#2a9d8f") +
  labs(
    title = "Early vs. Mid-Career Pay by School",
    x = "Early Career Pay (USD)",
    y = "Mid Career Pay (USD)") +
  geom_smooth(method = lm, se= F, color= "red" )
  theme_minimal()
```


```{r}
# stem vs mid career pay
ggplot(salary, aes(x = stem_percent, y = mid_career_pay)) +
  geom_point(alpha = 0.6, color = "#e76f51") +
  labs(
    title = "STEM Enrollment vs. Mid-Career Pay",
    x = "Percent in STEM",
    y = "Mid Career Pay (USD)") +
  geom_smooth(method = lm, se= F, color= "red" ) +
  theme_minimal()

```

```{r}
# state and early career pay
ggplot(salary, aes(x = reorder(state_name, early_career_pay, median), y = early_career_pay)) +
  geom_boxplot(fill = "#264653") +
  labs(
    title = "Early Career Pay by State",
    x = "State",
    y = "Early Career Pay (USD)"
  ) +
  coord_flip() +
  theme_minimal()
```

